### 2010-12-21 EquilibratedResults.tex
Reworded background, added new subsection to results

\section{Equilibrated MD}

\subsection{Water Orientation}

\subsection{\suldiox~Orientation}