### 2010-12-22 EquilibratedResults.tex
talking about how the surface location is determined

\section{Equilibrated MD}

Net molecular orientation was determined for the \wat~and \suldiox~molecules throughout the simulated system. Histograms were created to relate the angles $\theta$ and $\phi$ (see figure \ref{fig:water-angles}) to a molecule's position relative to the top surface of the aqueous slab.

\subsection{Water Orientation}

\subsection{\suldiox~Orientation}