### 2010-12-29 TransitResults.tex
New transit so2 angle analysis files - not yet added

\section{Steered MD}

\subsection {Neat-\wat~Single \suldiox System}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2transitangles.png}
		\caption{}
		\label{fig:so2-transit-angles}
	\end{center}
\end{figure}

\subsection {Saturated \wat~System}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2transitsaturatedangles.png}
		\caption{}
		\label{fig:so2-transit-saturated-angles}
	\end{center}
\end{figure}