### 2011-05-19 Article.tex
Rewrote the introduction to ask questions and emphasize the importance of the work, as per Geri's request

\documentclass{article}

\usepackage {amsmath}
\usepackage {setspace}
\usepackage [pdftex]{graphicx}
\usepackage[sort&compress]{achemso}
\usepackage{lineno}

% some formatting tags
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textwidth      6.5in

\title{Sulfur Dioxide on Water's Surface}
\author{Eric S. Shamay \and Kevin E. Johnson \and Geraldine L. Richmond}

\begin{document}

\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}

%\maketitle

%\onehalfspacing
\linenumbers 
\doublespacing

%\begin{abstract}
	%...Abstract goes here...
%\end{abstract}

\include{Introduction}
%\include{Background}
%\include{Methods}
%\include{DensityResults}
%\include{EquilibratedOrientation}
%\include{TransitOrientation}
%\include{Conclusions}

\bibliography{bib/SulfurDioxideText}

\end{document}