### 2011-05-19 DensityResults.tex
Rewrote the surface density results section to take out the ambiguity regarding the model used

\subsection{Surface Density Distributions}

One measure of surface activity is the spatial distribution of molecules in the interfacial region. The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations. The \suldiox~in the neat-water system remained at the top slab surface. The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained in the bulk.

%The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the top surface, $z$.

%\begin{equation}
%  \rho(z)=\frac{A}{\sqrt{2 \pi \sigma^2}} e^{-\frac{(z-\mu)^2}{2\sigma^2}}
%  \label{eq:gaussian}
%\end{equation}

In the simulated neat-water slab, the \suldiox~molecular density distribution indicates an affinity for the water surface. During the course of the simulation, the \suldiox~does not venture into the water bulk. Also, the \suldiox~does not escape the water surface into the gas phase, but remains located within 5-\angs~of the surface. This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS). The \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction. In a later section of this article we look at the effect of \suldiox~on surface water orientation.

The saturated surface simulation results indicate that under the high concentration conditions, \suldiox~accumulation at the surface is increased. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrate complex's geometry. However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods, and is furthermore verified by our experimental results.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density/density.png}
		\caption{Molecular density distributions of \wat~and \suldiox~calculated along the long-axis of the simulated cells. The position axis shows the distance of the given molecule from the top surface of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk. Distributions of both the neat-water simulation with a single \suldiox~(left, scaled 10x for clarity), and the saturated system (right) are shown.}
		\label{fig:density}
	\end{center}
\end{figure}