### 2011-06-28 Article.tex
Draft 7 and a new abstract!

\documentclass{article}

\usepackage {amsmath}
\usepackage {setspace}
\usepackage [pdftex]{graphicx}
\usepackage[sort&compress]{achemso}
\usepackage{lineno}

% some formatting tags
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textwidth      6.5in

\title{Sulfur Dioxide on Water's Surface}
\author{Eric S. Shamay \and Kevin E. Johnson \and Geraldine L. Richmond}

\begin{document}

\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}

%\maketitle

%\onehalfspacing
\linenumbers 
\doublespacing

\begin{abstract}
Sulfur dioxide binding to water surfaces is a model system for the study of gaseous adsorption processes on aqueous systems. Equilibrium and adsorbing \suldiox~molecules in a pure water interfacial system are modeled by classical molecular dynamics simulations using fully polarizable models for both \suldiox~and \wat. Neat-water systems with a single \suldiox, and high \suldiox~concentration saturated water systems are modeled at equilibrium and using steered molecular dynamics (SMD) to ascertain the orientational profile of a transiting \suldiox~molecule. Orientational depth profiling reveals the preferred orientations of \suldiox~during adsorption, and the complementary behaviors of surface waters adsorbing the gaseous solute.
\end{abstract}

\include{Introduction}
\include{Methods}
\include{Results}
\include{Conclusions}

\bibliography{bib/SulfurDioxideText}

\end{document}