### 2011-10-24 Article.tex
Polishing the review responses

\documentclass{article}

\usepackage {amsmath}
\usepackage {textcomp}
\usepackage {setspace}
\usepackage [pdftex]{graphicx}
\usepackage[sort&compress]{achemso}
\usepackage{lineno}

\usepackage[
top    = 0.5in,
bottom = 1.0in]{geometry}
%left   = 2.50cm,
%right  = 2.50cm]{geometry}

% some formatting tags
%\oddsidemargin  0.0in
%\evensidemargin 0.0in
%\textwidth      6.5in

\title{Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces}
\author{Eric S. Shamay \and Kevin E. Johnson \and Geraldine L. Richmond}

\begin{document}

\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}
%\newcommand{\deg}{$\,^{\circ}$}

\maketitle

%\onehalfspacing
\linenumbers 
\doublespacing

\begin{abstract}
One might expect the high surface tension of water to be a barrier to absorption of a gas into the liquid phase, but we know that gaseous adsorption onto, and subsequent absorption into a water surface is a common phenomenon on this planet.  What is not commonly known is how an atmospheric gas such as \suldiox~and molecules at the water surface can overcome the barrier created by strong water-water surface bonding interactions.  What this interplay looks like, the distances from the water surface at which these attractive interactions begin, and how they influence the orientational nature of both \suldiox~and surface water molecules is the focus of this computational study.  The results fill a void in the information about this system existing from previous experimental studies by providing information about the dimensional nature of the gas-surface interactions, and the details of how the two species twist and turn orientationally with increased surface interactions.  Classical molecular dynamics have been employed in both equilibrium and steered molecular dynamics (SMD) simulations for \suldiox~at a neat-water surface, and at a surface with high interfacial \suldiox~concentrations.  The results provide new molecular insights for understanding the interaction of this prevalent gas on aerosols and other aqueous surfaces in the environment.
\end{abstract}

\include{Introduction}
\include{Methods}
\include{Results}
\include{Conclusions}

\bibliography{bib/SulfurDioxideText}

\end{document}